---
title: "IS607 Week 3 Assignment"
author: "Anthony Pagan"
date: "September 12, 2018"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


####Read Data

```{r}
raw.data <-c("555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert")
```

####Recreate name with standard Firstnames and Lastnames

```{r}
library(stringr)
#original extracted names
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name

#Converted to Firstnames Lastnames 
names(name)<-c("v")
v<-str_split_fixed(name, ",", n=2)
t<-str_trim(sub(pattern = "\\w*\\.", replacement = "",x = str_c(v[,ncol(v)], v[,1], sep = " ")))
t
```

####Detect if name had a Title (ie.. Rev or DR)

```{r}
#List Names with Title
ttl <-unlist(str_extract_all(name, c("Rev.*","Dr.*")))
ttl

#State wether a name contains a Dr or Rev title
ttl.logic<-str_detect(name, c("Rev.*","Dr.*"))
ttl.log<-data.frame(t, ttl.logic)
names(ttl.log)<-c("Names", "Title")
ttl.log
```

####Detect if person had a middlename
```{r}
s.logic<-unlist(str_detect(name, " \\w{1}\\.{1}"))
s.log<-data.frame(t, s.logic)
names(s.log)<-c("Names", "MiddleName")
s.log
```

####Determine regular expression use case

This regular expression ([0-9]+\\\\$) can be used for any string that contains numbers followed by a $. Any other character can come before or  after, but has to contain a number followed by a $.

```{r}
stringz<- c("123$", "12333$$$","111a$", "aaa111$")
stringz.logic <-str_detect(stringz, "[0-9]+\\$")
data.frame(stringz, stringz.logic)

```

This regular expression (\\\\b[a-z]{1,4}\\\\b) would take up to 4 alphabet characters. If there is a 5 letters or any other letters than alpha it would be false.

```{r}
stringy<- c("dbddd", "abcd","111addd$", "aaaa")
stringy.logic <-str_detect(stringy, "\\b[a-z]{1,4}\\b")
data.frame(stringy, stringy.logic)

```

This regular expression (.*?\\\\.txt$) would be good for files with any characters and period and a txt at the end. This expression would be good find files that end in .txt.

```{r}
stringx<- c("dbddd.txt", "dddtxt","1114@.txt", ".txtaaaa")
stringx.logic <-str_detect(stringx, ".*?\\.txt$")
data.frame(stringx, stringx.logic)
```

This regular expression (\\\\d{2}/\\\\d{2}/\\\\d{4}) would for 2 digits followed by a slash, 2 digits followed by another slash and 4 digits. This would be good to detect a date if it includes a start and an end using ^ and $ like (^\\\\d{2}/\\\\d{2}/\\\\d{4}$)

```{r}
stringx<- c("dbdd22/2/2", "dd/22/22/2019dtxt","02/12/2009", "04/02/2017")
stringx.logic <-str_detect(stringx, "\\d{2}/\\d{2}/\\d{4}")
data.frame(stringx, stringx.logic)
stringxs.logic <-str_detect(stringx, "^\\d{2}/\\d{2}/\\d{4}$")
data.frame(stringx, stringxs.logic)
```

This regular expression (<(.+?)>.+?</\\\\1) would be good to use with html code.

```{r}
stringx<- c("<(a)>aa</(a)", "<(test)>aa</(test)","<(a)>aa", "<(a)>a")
stringx.logic <-str_detect(stringx, "<(.+?)>.+?</\\1")
data.frame(stringx, stringx.logic)
```

####Crack it with an R expression

```{r echo=TRUE}
j <-"clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5 fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

m<-str_extract_all(j, "[[:alnum:] ]*\\.?")
f<-str_c(str_extract_all(m, "[[:upper:]]"), sep = ".")
mf<- str_replace_all(f, "\" ?,? ?" , "")
str_replace_all(mf,"c\\(|\\)","")
```


