---
title: "Data 605 Assignment 11"
author: "Anthony Pagan"
date: "April 13, 2019"
output: 
      html_document:
        toc: false
        toc_float: true
        toc_depth: 3
        css: style.css
        highlight: pygments
        theme: cerulean
---

HW11

#Cars

Using the "cars" dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)


**Summary**

$\color{red}{\text{The residual lines appear to be normally distributed. The standard errors are at least 10 time smaller than the coefficient values. $\\$ This means there is little variability in the slope estimate. The very small Pr(>|t|) value shows the probability $\\$ that the values are NOT revelant. Result is they are revelant.}}$

```{r message=FALSE, warning=FALSE}
data("cars")

c.lm<-lm(cars$speed~cars$dist)

plot(cars$dist,cars$speed)
abline(c.lm)
summary(c.lm)

``` 


$\color{red}{\text{The fitted chart below shows the residuals uniformally scattered spread closely evenly throughout. $\\$ However, QQPlot shows that residuals are not normally distributed}}$

```{r}
plot(fitted(c.lm),resid(c.lm))
qqplot(fitted(c.lm),resid(c.lm))
qqline(fitted(c.lm),resid(c.lm))
```


