---
title: "Data 624 Homework 7"
author: "Anthony Pagan"
date: "4/5/2020"
output: 
    html_document:
        toc: true
        toc_float: true
        toc_depth: 5
        css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
#knitr::opts_chunk$set(echo = TRUE)
require(knitr)
library(ggplot2)
library(tidyr)
library(MASS)
library(psych)
library(kableExtra)
library(dplyr)
library(faraway)
library(gridExtra)
library(reshape2)
library(leaps)
library(pROC)
library(caret)
library(naniar)
library(pander)
library(pROC)
library(mlbench)
library(e1071)
library(fpp2)
library(mlr)
```


## 6.2. 

Developing a model to predict permeability (see Sect. 1.4) could save significant resources for a pharmaceutical company, while at the same time more rapidly identifying molecules that have a sufficient permeability to become a drug.

### (a) 

Start R and use these commands to load the data.The matrix fingerprints contains the 1,107 binary molecular predictors for the 165 compounds, while permeability contains permeability
response.

```{r}
library(AppliedPredictiveModeling)
data("permeability")
str( fingerprints)
#describe(fingerprints)
```

### (b) 

The fingerprint predictors indicate the presence or absence of substructures of a molecule and are often sparse meaning that relatively few of the molecules contain each substructure. Filter out the predictors that have low frequencies using the nearZeroVar function from the caret package. How many predictors are left for modeling?



```{r}
dim(fingerprints)
dim(permeability)
nzfinger<-nearZeroVar(fingerprints)
fingerprints <- fingerprints[,-nzfinger]
dim(fingerprints)

```

**After running nearZeroVar function there are `r dim(fingerprints)[2]` predictors left.**

### (c) 

Split the data into a training and a test set, pre-process the data, and
tune a PLS model. How many latent variables are optimal and what is
the corresponding resampled estimate of R2?

```{r}
require(pls)
library(AppliedPredictiveModeling)

ctrl<-trainControl(method="cv", number=10)
finger.df <- as.data.frame(fingerprints)
finger.df$permeability <- as.vector(permeability)
#trainp<- createDataPartition(finger.df, p=0.70)

set.seed(1)
trainp <- sample(1:nrow(finger.df), 0.7*nrow(finger.df))
trainf <- finger.df[trainp,]
testf <- finger.df[(nrow(trainf)+1):nrow(finger.df),]

plsfit <- plsr(permeability ~ ., data = trainf)
summary(plsfit)
plot(plsfit)

plsTune<-caret::train(trainf ,trainf$permeability,
                    method="pls",
                    tuneLength=20,
                    trControl=ctrl,
                    preProc = c("center", "scale"))
plot(plsTune)

```



### (d) 

Predict the response for the test set. What is the test set estimate of R2?

```{r}
pt<-predict(plsfit, testf)
head(pt)
summary(pt)

```

### (e) 

Try building other models discussed in this chapter. Do any have better predictive performance?

```{r}
require(MASS)
require(elasticnet)
require(caret)
require(lars)
enetfit <- enet(as.matrix(trainf), trainf$permeability, lambda = .01)
summary(enetfit)
plot(enetfit)
enetpred<-predict(enetfit, newx = testf, s=1, mode="fraction", type="fit")
summary(enetpred)
head(enetpred$fit)

enetGrid<-expand.grid(.lambda = c(0,.01, 1), .fraction = seq(.05, 1, length=20))
enetTune<-caret::train(trainf ,trainf$permeability,
                    method="enet",
                    tuneGrid=enetGrid,
                    trControl=ctrl,
                    preProc = c("center", "scale"))
plot(enetTune)

ridgeGrid<-data.frame(.lambda = seq(0, .1, length=15))
set.seed(100)
ridgeRegFit<- caret::train(trainf ,trainf$permeability,
                    method="ridge",
                    tuneGrid=ridgeGrid,
                    trControl=ctrl,
                    preProc = c("center", "scale"))
ridgeRegFit
summary(ridgeRegFit)
plot(ridgeRegFit)
ridgePred<-predict(ridgeRegFit, nex=testf)
ridgePred
```

### (f) 

Would you recommend any of your models to replace the permeability laboratory experiment?

**I would recommend using enet ridge regression **

## 6.3. 

A chemical manufacturing process for a pharmaceutical product was discussed in Sect. 1.4. In this problem, the objective is to understand the relationship between biological measurements of the raw materials (predictors), and the response of product yield. Biological predictors cannot be changed but can be used to assess the quality of the raw material before processing. On the other hand, manufacturing process predictors can be changed in the manufacturing process. Improving product yield by 1% will boost revenue by approximately one hundred thousand dollars per batch:

### (a) 

Start R and use these commands to load the data. The matrix processPredictors contains the 57 predictors (12 describing the input biological material and 45 describing the process predictors) for the 176 manufacturing runs. yield contains the percent yield for each run.

```{r}
library(AppliedPredictiveModeling)
data("ChemicalManufacturingProcess")
cm<-ChemicalManufacturingProcess
dim(cm)

```

### (b) 

A small percentage of cells in the predictor set contain missing values. Use
an imputation function to fill in these missing values (e.g., see Sect. 3.8).

```{r}

cm<-cm[complete.cases(cm),]
anyNA(cm)
```

### (c) 

Split the data into a training and a test set, pre-process the data, and tune a model of your choice from this chapter. What is the optimal value of the performance metric?

```{r}

set.seed(1)
trainp <- sample(1:nrow(cm), 0.7*nrow(cm))
trainf <- cm[trainp,]
testf <- cm[(nrow(trainf)+1):nrow(cm),]

plsfit <- plsr(Yield ~ ., data = cm)
summary(plsfit)
plot(plsfit)

```

### (d) 

Predict the response for the test set.What is the value of the performance metric and how does this compare with the resampled performance metric on the training set?

```{r}
library(caret)
trctrl<- trainControl(method="repeatedcv", number=10,repeats=3)
svm_lin<- caret::train(Yield~., data=cm, method="svmLinear", 
                trControl=trctrl, preProcess=c("center","scale"),
                tuneLength =10)
svm_lin
test_pred<- predict(svm_lin, newdata=testf)
test_pred

```

### (e) 

Which predictors are most important in the model you have trained? Do either the biological or process predictors dominate the list?

```{r}
plot(test_pred)
```

### (f) 

Explore the relationships between each of the top predictors and the response. How could this information be helpful in improving yield in future runs of the manufacturing process?



## APPENDIX

**Code used in analysis**
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}

```

