---
title: "Data 605 Assignment 8"
author: "Anthony Pagan"
date: "March 30, 2019"
output: 
      html_document:
        toc: true
        toc_float: true
        toc_depth: 3
        css: style.css
        highlight: pygments
        theme: cerulean
---

HW9

#Stocks

The price of one share of stock in the Pilsdorff Beer Company (see Exercise
8.2.12) is given by Yn on the nth day of the year. Finn observes that
the differences Xn = Yn+1 ??? Yn appear to be independent random variables
with a common distribution having mean ?? = 0 and variance 2 = 1/4. If
Y1 = 100, estimate the probability that Y365 is

##A) >=100

```{r message=FALSE, warning=FALSE}
a<-f(2)
```     

**``**

##B)  >=110

```{r message=FALSE, warning=FALSE}
b<-f(5)
```     

**``**

##C) >=120

```{r message=FALSE, warning=FALSE}
c<-f(9)
```     

**``**

#Binomial Distribution

Calculate the expected value and variance of the binomial distribution using the moment generating function.

```{r message=FALSE, warning=FALSE}
d<-f(20)
```     

**``**

#Exponential Distribution

Calculate the expected value and variance of the exponential distribution using the moment generating function.


```{r message=FALSE, warning=FALSE}
d<-f(20)
```     

**``**