---
title: "Data 605 Assignment 10"
author: "Anthony Pagan"
date: "April 06, 2019"
output: 
      html_document:
        toc: false
        toc_float: true
        toc_depth: 3
        css: style.css
        highlight: pygments
        theme: cerulean
---

HW10

#Bets

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.
A guard agrees to make a series of bets with him. If Smith bets A dollars,
he wins A dollars with probability .4 and loses A dollars with probability .6.
Find the probability that he wins 8 dollars before losing all of his money if:

Equations:

* N Number of Wins: ${\frac{1-(\frac{q}{p})^i}{1-(\frac{q}{p})^N}}$
* Inf Win= 1 - ${(\frac{q}{p})^i}$
* Losing all = 1 - ${(\frac{q}{p})}$

```{r message=FALSE, warning=FALSE}

f<-function(i)
{
p <- .6
q <- .4
N<-8
iw <- round((1-(q/p)^i),2)
nw <- round((iw/(1-(q/p)^N)),2)
l<- 1-iw
pa<-c(iw,nw,l)
}

``` 

```{r message=FALSE, warning=FALSE}
a<-f(1)
```     


**Bets 1 dollar each time (timid strategy).**

$\color{red}{\text{The probability he wins \$8 before losing if he bets \$1 is `r a[2]`}}$


**He bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy).**

$\color{red}{\text{The probability he wins \$8 before losing if he bets as much as possible is `r a[1]`}}$

**Which strategy gives Smith the better chance of getting out of jail?**

$\color{red}{\text{The best strategy is to bet \$1 at a time. Below shows probability of winning \$8 is slighly higher per .01}}$

```{r}
sapply(seq(0,1, by=.01), f)
```

