---
title: "Data 605 Assignment 3"
author: "Anthony Pagan"
date: "February 16, 2019"
output: html_document
---

(1) What is the rank of the matrix A? `r matrix(c(1,-1,0,5,2,0,1,4,3,1,-2,-2,4,3,1,3),4,4,byrow=FALSE)`

 
#### The Matix Rank is the number of non zero rows in the reduce row echelon of a Matrix

```{r message=FALSE, warning=FALSE}
library(matlib)
A<- matrix(c(1,-1,0,5,2,0,1,4,3,1,-2,-2,4,3,1,3),4,4,byrow=FALSE)
A
echelon(A, reduce=FALSE,verbose=TRUE)
A.rank<-R(A)

```

####In this example the rank of matrix A is `r A.rank`


(2) Given an mxn matrix where m > n, what can be the maximum rank? The mini-
mum rank, assuming that the matrix is non-zero?

####The Maximum rank for an mxn matrix number of linearly independent non-zero m columns in the matrix = n when n < m or = m when n > m. The minimum rank will be the min(m, n). If m<n min rank is m, if n<m min rank is n.

(3) What is the rank of matrix B?
B = `r matrix(c(1,2,1,3,6,3,2,4,2),3,3,byrow=TRUE)`

```{r}
A<- matrix(c(1,2,1,3,6,3,2,4,2),3,3,byrow=TRUE)
A
echelon(A, reduce=FALSE)
A.rank<-R(A)
```

####In this example the rank of matrix A is `r A.rank`

Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your
work. You'll need to write out the characteristic polynomial and show your solution.

A = `r matrix(c(1,2,3,0,4,5,0,0,6),3,3,byrow=TRUE)`


```{r}
#The Matrix
A<- matrix(c(1,2,3,0,4,5,0,0,6),3,3,byrow=TRUE)
A
```

####The eigen values of A are:

```{r}
#Get eigen values and vectors

e<-eigen(A)
e.values<-e$value
e.values
```

####The eigen vectors of A are:

```{r}
e.vectors<-e$vector
e.vectors

```


####The polynomial equation derived from eigen values is:

`r -e.values[1]`$\mathbf\lambda^3$\ - (`r -e.values[2]`) $\mathbf\lambda^2$\ + (`r -e.values[3]`) $\mathbf\lambda$\ - `r det(A)`


####Testing Eigen values Null Space

```{r}

#Create identity matrix 
B<-matrix(c(1,0,0,0,1,0,0,0,1),3,3,byrow=TRUE)
B

#get determinate of  Identity matrix minus A 
(A[1,1]*(A[2,2]*A[3,3]))-(A[2,1]*(A[1,3]*A[3,3]))+(A[1,3]*(A[2,1]*A[3,2]))
det(B-A)
det(A)
B-A
#Test eigen values

#Test first eigen value minus A
C1<-A
C1[1,1]<-e.values[1]-C1[1,1]
C1[2,2]<-e.values[1]-C1[2,2]
C1[3,3]<-e.values[1]-C1[3,3]
C1

#Reduced row echelon of eigen vector
C1.e<-echelon(C1, reduced =FALSE)
C1.e

#Eigen value minus vector A to confirm determinant equals 0
C1.v<-C1-A
C1.v

#Test vector multiplied by eigen vector values

C1.v*C1.e - C1.v

#Test second eigen value minus A
C2<-A
C2[1,1]<-e.values[2]-C2[1,1]
C2[2,2]<-e.values[2]-C2[2,2]
C2[3,3]<-e.values[2]-C2[3,3]
C2

#Reduced row echelon of eigen vector
C2.e<-echelon(C1, reduced =FALSE)
C2.e
#Eigen value minus vector A to confirm determinant equals 0
C2.v<-C2-A
C2.v
#Test vector multiplied by eigen vector values

C2.v*C2.e - C2.v

#Test third eigen value minus A
C3<-A
C3[1,1]<-e.values[3]-C3[1,1]
C3[2,2]<-e.values[3]-C3[2,2]
C3[3,3]<-e.values[3]-C3[3,3]
C3

#Reduced row echelon of eigen vector
C3.e<-echelon(C1, reduced =FALSE)
C3.e
#Eigen value minus vector A to confirm determinant equals 0
C3.v<-C3-A
C3.v
#Test vector multiplied by eigen vector values

C3.v*C3.e - C3.v

```

