---
title: "Data 605 Assignment 15"
author: "Anthony Pagan"
date: "May 09 2019"
output: 
      html_document:
        toc: true
        toc_float: true
        toc_depth: 3
        css: style.css
        highlight: pygments
        theme: cerulean
---

HW15

#Regression

Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.

( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )

```{r}
x <-c(5.6,6.3,7,7.7,8.4)
y<-c(8.8,12.4,14.8,18.2,20.8)
l<-lm(y~x)
plot(x,y)
abline(l)
```

**Regression Line: Y = `r l$coefficients[1]` + `r round(l$coefficients[2],2)`x**

#Maxima/Minima

Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form( x, y, z ). Separate multiple points with a comma.

${f(x,y)=24x-6xy^2-8y^3}$

 * ${f(x)= x -6y^2= 0}$ = ${x = 6y^2}$
 * ${f(y)= 12y - 24y = 0}$ = ${y = -12y = \frac{-1}{12}}$

```{r}
y <--1/12
x <- 6*y^2
z <- 12*x-6*x*y^2-8*y^3
a <- c(x,y,z)
```

**The local maxima, minima, and saddle points are (`r a`)**

#Functions

A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell 81 - 21x + 17y units of the "house" brand and 40 + 11x - 23y units of the "name" brand.

Step 1. Find the revenue function R ( x, y ).

```{r}
uh <- 81 - 21*x + 17*y
un <- 40 + 11*x - 23*y

```

**Revenue function = (81 - 21*x + 17*y) + (40 + 11*x - 23*y) **

Step 2. What is the revenue if she sells the "house" brand for $2.30 and the "name" brand for $4.10?

```{r}

x <- 2.30
y <- 4.10
R <- uh+un
```

**If she sells the "house" brand for $2.30 and the "name" brand for $4.10 the revenue will be $`r round((R),2)`**

#Minimize Cost

A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a product each week. The total weekly cost is given by ${C(x, y) = \frac{1}{6}x^2 + \frac{1}{6}y^2+ 7x + 25y + 700}$, where x is the number of units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

 * ${f(x)= 2\frac{1}{6}x+7 = 0}$ 
 * ${f(y)= 2\frac{1}{6}x+25 = 0}$

```{r}
x<-7/(2/6)
y<-25/(2/6)

```

**To minmize weekly cost `r x` units should be produced in Los Angeles and `r y` units should be producded in Denver**

#Integral

Evaluate the double integral on the given region.

${\int\int_R e^{8x+3y}dA:R:2\le{x}\le{4}}$ and ${2\le{y}\le{4}}$

Write your answer in exact form without decimals.

```{r}
f<-function(x) { exp(8*x+3*y) }
f2<- function(y) { sapply(y, 
function(z) { integrate(f, 2, 4)$value }) }
v<- integrate(f2 , 2, 4)
```

**value is `r v$value`**
