---
title: "Data 605 Assignment 8"
author: "Anthony Pagan"
date: "March 30, 2019"
output: 
      html_document:
        toc: true
        toc_float: true
        toc_depth: 3
        css: style.css
        highlight: pygments
        theme: cerulean
---

HW9

#Stocks

The price of one share of stock in the Pilsdorff Beer Company (see Exercise
8.2.12) is given by Yn on the nth day of the year. Finn observes that
the differences Xn = Yn+1 - Yn appear to be independent random variables
with a common distribution having ${\mu}$ = 0 and variance ${\sigma^2}$ = 1/4. If
Y1 = 100, estimate the probability that Y365 is

```{r message=FALSE, warning=FALSE}

f<-function(y2)
{
m <- 0
s <- .25
y<-100
x <- 365
s2 = x*sqrt(s)
z<- (y-y2)/s2
pnorm(z)
}

``` 

##A) >=100

```{r message=FALSE, warning=FALSE}
a<-f(100)
```     


**The probablilty of Y365 >= 100 is `r round(a, 2)`**

##B)  >=110

```{r message=FALSE, warning=FALSE}
b<-f(110)
```     

**The probablilty of Y365 >= 110 is `r round(b, 2)`**

##C) >=120

```{r message=FALSE, warning=FALSE}
c<-f(120)
```     

**The probablilty of Y365 >= 120 is `r round(c, 2)`**

#Binomial Distribution

Calculate the expected value and variance of the binomial distribution using the moment generating function.

Binomial MGF = ${[pe^t+ (1-p)]^n}$

```{r message=FALSE, warning=FALSE}
amgf <- (a*(log(exp(365))+ (1-a)))^1
bmgf <- (b*(log(exp(365))+ (1-b)))^1
cmgf <- (c*(log(exp(365))+ (1-c)))^1
mmgf <- mean(c(amgf,bmgf,cmgf))
smgf <- var(c(amgf,bmgf,cmgf))
```     

**The expected value of the binomial distirbution is `r round(mmgf,2)` and the variance is `r round(smgf, 2)`**

#Exponential Distribution

Calculate the expected value and variance of the exponential distribution using the moment generating function.

* Exponential MGF = ${\frac{\lambda}{\lambda-t}}$
* E[X]= ${\frac{1}{\lambda}}$
* Var[X]= ${\frac{1}{\lambda^2}}$

```{r message=FALSE, warning=FALSE}
eamgf<-1/a*100
ebmgf<-1/b*110
ecmgf<-1/c*120
emmgf<-mean(c(eamgf,ebmgf,ecmgf))

samgf<-((100-eamgf)^2)/100
sbmgf<-((110-ebmgf)^2)/110
scmgf<-((120-ecmgf)^2)/120
smmgf<-mean(c(samgf,sbmgf,scmgf))
```     

**The expected value of the exponential distribution is `r round(emmgf,2)` and the variance is `r round(smmgf, 2)`**