---
title: "Data 605 Assignment 13"
author: "Anthony Pagan"
date: "April 29, 2019"
output: 
      html_document:
        toc: true
        toc_float: true
        toc_depth: 3
        css: style.css
        highlight: pygments
        theme: cerulean
---

HW13

#1) Integration by substitution

Use integration by substitution to solve the integral below.

${\int{4e^{-7x}dx}}$ = 

 * u = ${e^{-7x}}$
 * ${{4}\int e^{-7x}dx}$

#2) Functions
Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of ${\frac{dN}{dt}=\frac{3150}{t^4}-2}$ bacteria per cubic centimeter per day, where t is the number of days since treatment began. Find a function N( t ) to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter.

f(x)= ${{6530}{t^{1/4}}-2}$ =

 * ${\int_1^tf(x)\frac{dN}{dt}}$


#3) Area Functions
Find the total area of the red rectangles in the figure below, where the equation of the line is f ( x ) = 2x-9

 * f(x) = 2x -9
 * Solve for y
 * 2x - 9 = 0
 * x = 4.5
 * y = 2x -9
 * A = 1/2(x*y)
```{r}
integrand <- function(x) {2*x-9} 
A<-integrate(integrand, lower = 4, upper = 9)
A<-A[1]
```


**Total area is approx `r A`**

#4) Graph Areas
Find the area of the region bounded by the graphs of the given equations.
y = x2-2x-2, y = x + 2


```{r}
eq <- function(x){x^2-2*x-2}
eq2<-function(x){x+2}
plot(eq(1:5), type='l')
lines(eq2(1:5), type='l')

integrand <- function(x) {(x^2-2*x-2)-(x+2)} 
A<-integrate(integrand, lower = 1, upper = 4)
A<-A[1]
```

**Area of region is approx `r A`**

#5) Word Problems
A beauty supply store expects to sell 110 flat irons during the next year. It costs $3.75 to store one flat iron for one year.There is a fixed cost of $8.25 for each order. Find the lot size and the number of orders per year that will minimize inventory costs.

* Expected Unit Sales = 110
* Storage Costs = 3.75*(110-UnitsSold)
* Fixed Costs = -x*8.25

```{r}
x<-110
eq<- function(x){3.75*(110-x)+x*8.25}
plot(eq(1:100), type='l')
integrand <- function(x) {3.75*(110-x)+x*8.25}
A<-integrate(integrand, lower = 0, upper = 1)
A<-as.double(A[1])
A<-A/100
```

**Selling `r A` lots of a 100 will minimize cost**

#6) Integration II
Use integration by parts to solve the integral below.
${\int{ln(9x)}* x^6dx}$

* u = ln(9x) = ${\frac{9}{x}}$
* v = ${x^6dx}$ = ${5x^5}$
* = -ln(${\frac{9}{x}}$)-${5x^5}$+C

#7) Density Function
Determine whether f ( x ) is a probability density function on the interval ${[1,e^6}]$ . If not, determine the value of the definite integral.

f(x) = ${\frac{1}{6x}}$ =

* ${\int_0^1\frac{-1}{6x^2}dx}$

```{r}
eq<-function(x){1/(6*x)}
eq2<- function(x){exp(x)}
plot(eq(1:100), type='l')
plot(eq2(1:100),type='l')

integrand <- function(x) {-1/(6*x^2)} 
A<-integrate(integrand, lower = 1, upper = 10)
A<-A[1]
E<-integrand(1)
O<-integrand(exp(6))

```


**The integrate is `r A`. The results when x = ${e^6}$ is `r E` and the results for when x =1 is `r O`. The probability is `r as.double(A)-as.double(E[1])+(O[1])/as.double(E[1])+as.double(O[1])`**