---
title: "Data 605 Assignment 7"
author: "Anthony Pagan"
date: "March 16, 2019"
output: 
      html_document:
        toc: true
        toc_float: true
        toc_depth: 3
        css: style.css
        highlight: pygments
        theme: cerulean
---

HW7

#1) Uniform Distribution

Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi's. Find the distribution of Y .

**The uniform distribution formula of Y is Y = 1/xa-xa**

```{r message=FALSE, warning=FALSE}
a <- runif(1000)
d <-density(1/(max(a)-min(a)),min(a))
plot(d)
d
```     

#2) Machine Lifetime

Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer's expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).

##a) Geometric

What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years..)

```{r message=FALSE, warning=FALSE}
c<-c(1,2,3,4,5,6,7,8,9,10)
f<- 1/10
nf<- 9/10
n <- 8
m<- 1/10
nf8<- round(nf^n,2)
ev<-sum(c[1:8]*nf+c[9:10]*f)/10
std<-sd(c[1:8]*nf+c[9:10]*f)
```

**The geometric probability of the machine failing after 8 years is `r nf8`, expected value is `r ev` and standard deviation is `r std`**

##b) Exponential

What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.

```{r message=FALSE, warning=FALSE}
nf8 <- round(exp(log(nf^n)),2)
ev<-sum(c[1:8]*nf+c[9:10]*f)/10
std<-sd(c[1:8]*nf+c[9:10]*f)
```

**The expected exponential probability of the machine failing after 8 years is `r nf8`**

##c) Binomial

What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)

```{r message=FALSE, warning=FALSE}
f<-dbinom(8,10,.9)
nf<- 1-dbinom(8,10,.9)
cf<- c(9,10)
cnf<-c(1,2,3,4,5,6,7,8)
ecf <- sum(f*cf)/10
ecnf <- sum(nf*cnf)/10
sdcf <- sd(f*cf)
sdnf <- sd(nf*cnf)
ef<- round(ecf + ecnf, 2)
std<- round(sdcf + sdnf, 2)
nf8 <- round((1- f*nf)/2, 2)
```

**The expected  probability of the machine as a binomial failing after 8 years is `r nf8`, standard deviation is `r std` and expected value is `r ef`**

##d) Poisson


What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson.

```{r message=FALSE, warning=FALSE}
f<-rpois(8, .1)
nf<-rpois(8, .9)
cf<- c(9,10)
cnf<-c(1,2,3,4,5,6,7,8)
ecf <- sum(f*cf)/10
ecnf <- sum(nf*cnf)/10
sdcf <- sd(f*cf)
sdnf <- sd(nf*cnf)
ref<- round(ecf + ecnf, 2)/2
rstd<- round(sdcf + sdnf, 2)/2
rnf8 <- round((1-(sum(f)/length(f))*(sum(nf)/length(nf))), 2)/2

```

**The expected  probability and SD of the machine failing after 8 years is `r rnf8`, standard deviation is `r rstd` and expected value is `r ref`**