---
title: "Data 605 Assignment 5"
author: "Anthony Pagan"
date: "March 1, 2019"
output: 
      html_document:
        toc: true
        toc_float: true
        toc_depth: 3
        css: style.css
---

Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. Prove that B and C are proper probability distributions.
Note that the point (B,C) is then chosen at random in the unit square.
Find the probability that

#(a) B + C < 1/2

```{r}
B=seq(0,1,length=2000)
C=seq(0,1,length=2000)
y=dnorm(punif(B, min = 0, max = 1) > runif(C, min = 0, max = 1))
boxplot(subset(y,y<.5),type="l", lwd=2, col="blue")
length(subset(y,y<.5))/length(y)
hist(y)

```

#(b) BC < 1/2

```{r}
B=seq(0,1,length=2000)
C=seq(0,1,length=2000)
y=dnorm(punif(B, min = 0, max = 1) * runif(C, min = 0, max = 1))
boxplot(subset(y,y<.5),type="l", lwd=2, col="blue")
length(subset(y,y<.5))/length(y)
hist(y)
```

#(c) |B - C| < 1/2

```{r}
B=seq(0,1,length=2000)
C=seq(0,1,length=2000)
y=dnorm(punif(B, min = 0, max = 1) - runif(C, min = 0, max = 1))
boxplot(subset(y,y<.5),type="l", lwd=2, col="blue")
length(subset(y,y<.5))/length(y)
hist(y)
```

#(d) max{B,C} < 1/2

```{r}
B=seq(0,1,length=2000)
C=seq(0,1,length=2000)
y=dnorm(max(punif(B, min = 0, max = 1), runif(C, min = 0, max = 1)))
boxplot(subset(y,y<.5),type="l", lwd=2, col="blue")
length(subset(y,y<.5))/length(y)
```

#(e) min{B,C} < 1/2

```{r}
B=seq(0,1,length=2000)
C=seq(0,1,length=2000)
y=dnorm(min(punif(B, min = 0, max = 1) , runif(C, min = 0, max = 1)))
boxplot(subset(y,y<.5),type="l", lwd=2, col="blue")
length(subset(y,y<.5))/length(y)
```