---
title: "Data 605 Assignment 6"
author: "Anthony Pagan"
date: "March 9, 2019"
output: 
      html_document:
        toc: true
        toc_float: true
        toc_depth: 3
        css: style.css
        highlight: pygments
        theme: cerulean
---

HW6

#1) Box of Marbles
A box contains 54 red marbles, 9 white marbles, and 75 blue marbles. If a marble is randomly selected from the box,
what is the probability that it is red or blue? Express your answer as a fraction or a decimal number rounded to four
decimal places.

```{r}

rm <- 54
wm <- 9
bm <- 75
am <- sum(rm,wm , bm)

rbm <- round((rm/am + bm/am),4) 

```

**The probability that marbles will be red or blue is `r rbm`**

#2) Golf Balls
You are going to play mini golf. A ball machine that contains 19 green golf balls, 20 red golf balls, 24 blue golf balls, and
17 yellow golf balls, randomly gives you your ball. What is the probability that you end up with a red golf ball? Express
your answer as a simplified fraction or a decimal rounded to four decimal places.

```{r}
gb <- 19
rb <- 20
bb <- 24
yb <- 17
ab <- sum(gb, rb, bb, yb)
prb <- round((rb/ab),4)
```

**The probability of getting a red golf ball is `r prb`**

#3) Pizza Delivery

A pizza delivery company classifies its customers by gender and location of residence. The research department has
gathered data from a random sample of 1399 customers. The data is summarized in the table below.
Gender and Residence of Customers
Males Females
Apartment 81 228
Dorm 116 79
With Parent(s) 215 252
Sorority/Fraternity House 130 97
Other 129 72

What is the probability that a customer is not male or does not live with parents? Write your answer as a fraction or a
decimal number rounded to four decimal places.
Determine if the following events are independent.
Going to the gym. Losing weight.

```{r}
f <-c(228,79,252,97,72)
m <-c(81,116,215,130,129)
lwp <-c(215,252)
sf <-sum(f)
sm <-sum(m)
stotal <-sum(sf,sm)
slwp <- sum(lwp)
nm <- round((1 -sm/stotal),4)
snlwp <- round((1 - slwp/stotal),4)
nmsnlwp<- nm+snlwp - round((lwp[1]/slwp),4)

```

**The probability customer is not male or does not live with parents is `r nmsnlwp`**

#4) Gym 

Determine if the following events are independent.
Going to the gym. Losing weight.
Answer: A) Dependent B) Independent

**Going to the gym and losing weight are independent. Not all people go to gym to lose weight. Some go to the gym to gain muscle mass and/or just to stay healthy. Also, some people who go to the gym to lose weight don't always succeed.**

#5) Veggie Wrap
A veggie wrap at City Subs is composed of 3 different vegetables and 3 different condiments wrapped up in a tortilla. If
there are 8 vegetables, 7 condiments, and 3 types of tortilla available, how many different veggie wraps can be made?

```{r}
v <- 8
c <- 7
t <- 3
vw <- (factorial(v)/(factorial(3) * factorial(v-3))) * (factorial(c)/(factorial(3) * factorial(c-3))) * t
choose(8,3)
```

 **`r vw` different veggie wraps can be made**

#6) Gas and News
Determine if the following events are independent.
Jeff runs out of gas on the way to work. Liz watches the evening news.

Answer: A) Dependent B) Independent

**These events are independent. Running out of gas has no dependency with Liz watching the eveninng news.**

#7) Candidates
The newly elected president needs to decide the remaining 8 spots available in the cabinet he/she is appointing. If there
are 14 eligible candidates for these positions (where rank matters), how many different ways can the members of the
cabinet be appointed?

```{r}
s <- 8
c <- 14

a<- factorial(s) / factorial(c-s)

```

**There are `r a` different ways to appoint members to the cabinet.**

#8) Jellybeans
A bag contains 9 red, 4 orange, and 9 green jellybeans. What is the probability of reaching into the bag and randomly
withdrawing 4 jellybeans such that the number of red ones is 0, the number of orange ones is 1, and the number of green
ones is 3? Write your answer as a fraction or a decimal number rounded to four decimal places.

```{r}
r <- 9
o <- 4
g <- 9

a<- sum(r,o,g)
pr <- round(r/a,4)
po <- round(o/a,4)
pg <- round(g/a,4)

prob<-round(((1-pr)*po*(3*pg)),4)
```

**The probability of withrawing 4 jellybeans where 0 are red, 1 is orange and 3 are green is `r prob`**

#9) Evaluate Expression
Evaluate the following expression.
11!
7!

```{r}
f<- factorial(11) / factorial(7)
```


**The expression gives the number of different permutations if 7 items were select from 11 where order does not matter. In this case the number of permuations are `r f`**

#10) The complement

Describe the complement of the given event.
67% of subscribers to a fitness magazine are over the age of 34.

**The complement of 67% of subscribers of a fitness magazine over age 34 would be 100-67% or 33% of subscribers that are under the age of 34**

#11) Coin Toss
If you throw exactly three heads in four tosses of a coin you win $97. If not, you pay me $30.
Step 1. Find the expected value of the proposition. Round your answer to two decimal places.
Step 2. If you played this game 559 times how much would you expect to win or lose? (Losses must be entered as
negative.)

```{r}
#Step 1
h <- dbinom(3,4,.5)
t <- dbinom(1,4,.5)

w <- 97
l <- -30
wev <- h*w
lev <- t*l
exp1 <- round((wev + lev),2)

#Step 2
h <- mean(rbinom(559,4,.5)==3)
t <- mean(rbinom(559,4,.5)==1)

wev <- h*w
lev <- t*l
exp2 <- round((wev + lev),2)
```

**STEP1 The expected value of 4 coins tosses is `r exp1`**
**STEP2 The expected value of 559 coin tosses is `r exp2`**

#12) Coin Toss 2

Flip a coin 9 times. If you get 4 tails or less, I will pay you $23. Otherwise you pay me $26.
Step 1. Find the expected value of the proposition. Round your answer to two decimal places.
Step 2. If you played this game 994 times how much would you expect to win or lose? (Losses must be entered as
negative.)

```{r}
#Step 1
h <- dbinom(4,9,.5)
t <- dbinom(5,9,.5)

w <- 23
l <- -26
wev <- h*w
lev <- t*l
exp1 <- round((wev + lev),2)

#Step 2
h <- mean(rbinom(994,9,.5)<=4)
t <- mean(rbinom(994,9,.5)>4)

wev <- h*w
lev <- t*l
exp2 <- round((wev + lev),2)
```

**STEP1 The expected value of 9 coins tosses is `r exp1`**
**STEP2 The expected value of 994 coin tosses is `r exp2`**

#13) Polygraph

The sensitivity and specificity of the polygraph has been a subject of study and debate for years. A 2001 study of the use of polygraph for screening purposes suggested that the probability of detecting a liar was .59 (sensitivity) and that the probability of detecting a "truth teller" was .90 (specificity). We estimate that about 20% of individuals selected for the screening polygraph will lie.

**The below matrix provides the probability matrix for answers a-c**

```{r}

l <- .59
t <- .9
ll <- .2
tt <- .8

matrix(c(ll,l, ll*l, ll, 1-l, ll*(1-l), tt, t, t*tt, tt,  1-t, tt*(1-t)), 4,3,byrow =TRUE)
```

##a. Liar
What is the probability that an individual is actually a liar given that the polygraph detected him/her as such? (Show me the table or the formulaic solution or both.)


```{r}

lie<- l*ll

```

**The probabiliy that the individual is a liar given they are detected as such is `r lie`**

##b. Truth-teller
What is the probability that an individual is actually a truth-teller given that the polygraph detected him/her as such? (Show me the table or the formulaic solution or both.)

```{r}
truth<- tt*t

```


**The probabiliy that the individual is a true-teller given they are detected as such is `r truth`**

##c. Liar-Liar
What is the probability that a randomly selected individual is either a liar or was identified as a liar by the polygraph? Be sure to write the probability statement.

```{r}
lil<- lie + (tt*(1-t))
```

**The probabiliy that the individual is either a liar or they are detected as such is `r lil`**