---
title: "IS607 Project 2"
author: "Anthony Pagan"
date: "October 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####Load Required Libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
##Load libraries
library(dplyr) #all data
library(tidyr) #all data
library(readxl)#Import data 2
library(ggplot2)#all data
library(plotrix) #data1
library(plotly) #data 2 https://plot.ly/ggplot2/animations/
library(gapminder) #data2 https://plot.ly/ggplot2/animations/
```

#####Data1 posted by Michael Silva (Adult Arrests)
##### https://data.ny.gov/Public-Safety/Adult-Arrests-by-County-Beginning-1970/rikd-mt35/data. 

####Get Data 1

```{r data1, message=FALSE, warning=FALSE, paged.print=TRUE}
theurl <- "https://data.ny.gov/api/views/rikd-mt35/rows.csv?accessType=DOWNLOAD"
thedata <- read.table(file = theurl, header = TRUE, sep = ",")
glimpse(thedata)
```

####Tidy Data Offense
Gather all offense totals and display GGPlots

```{r message=FALSE, warning=FALSE}
head(thedata)
tddata1.1<-thedata%>%
    select( -Felony.Total, -Misdemeanor.Total, -Year, -Total) %>%
    gather(offense, value, 2:9)%>%
    group_by(offense)%>%
    summarise_each(funs(sum), value)%>%
    mutate(pctallcrime = round(value/sum(thedata$Total)*100,5))%>%
    arrange(desc(value))
tddata1.1
```

####GGPlot

```{r message=FALSE, warning=FALSE}
options("scipen" = 20)
lbls <- paste(tddata1.1$offense, round(tddata1.1$pctallcrime,0)) 
lbls <- paste(lbls,"%",sep="") 
pie(tddata1.1$value, labels = lbls, col = rainbow(length(tddata1.1$value)),  main="Pie Chart of Offenses")
 
ggplot(data= tddata1.1) +
    geom_point(mapping = aes(x = pctallcrime, y = value, size = pctallcrime, color = offense))

ggplot(data= tddata1.1) +
    geom_bar(mapping = aes(x = offense, y = value, fill= offense), stat = "identity", position = "identity")+
#theme(axis.text.x=element_text(angle=-90))
theme(axis.text.x = element_blank())
```

####Tidy Data by Top 10 Counties
Gather all offense totals and display GGplots by Counties
```{r message=FALSE, warning=FALSE}
tddata1.2<-thedata%>%
    select( -Felony.Total, -Misdemeanor.Total, -Year, -Total) %>%
    gather(offense, value, 2:9)%>%
    group_by(County)%>%
    summarise_each(funs(sum), value)%>%
    mutate(pctallcrime = round(value/sum(thedata$Total)*100,5))%>%
    arrange(desc(value))%>%
    top_n(10, value)
tddata1.2
```

####GGPlot

```{r message=FALSE, warning=FALSE}
lbls <- paste(tddata1.2$County, round(tddata1.1$pctallcrime,0)) 
lbls <- paste(lbls,"%",sep="") 
pie(tddata1.2$value, labels = lbls, col = rainbow(length(tddata1.2$value)),  main="Pie Chart by Top 10 County")

ggplot(data= tddata1.2) +
    geom_point(mapping = aes(x = pctallcrime, y = value, size = pctallcrime, color = County))
```


```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=6}
tddata1.3<-thedata%>%
    select( -Felony.Total, -Misdemeanor.Total, -Year, -Total) %>%
    gather(offense, value, 2:9)%>%
    group_by(County, offense)%>%
    summarise_each(funs(sum), value)%>%
    mutate(pctallcrime = round(value/sum(thedata$Total)*100,5))%>%
    arrange(desc(value))

ggplot(data= tddata1.3) +
    geom_bar(mapping = aes(x = County, y = value, fill= offense), stat = "identity", position = "identity")+
    theme(axis.text.x=element_text(angle=-90, size=9))
```

####Data2 posted by Juanelle Marks (Population Migration)
#### http://www.un.org/en/development/desa/population/migration/data/estimates2/estimates17.shtml
 
####Get Data 2

```{r data2, message=FALSE, warning=FALSE}
theurl2 <- "http://www.un.org/en/development/desa/population/migration/data/empirical2/data/UN_MigFlow_Totals.xlsx"
destfile <- "UN_MigFlow_Totals.xlsx"
download.file(theurl2, destfile, mode = "wb")
thedata2<-read_xlsx("UN_MigFlow_Totals.xlsx", skip = 16)
glimpse(thedata2)
```

####Tidy Data by Year Totals
Gather all year totals and display dynamic year to year changes in plotly

```{r message=FALSE, warning=FALSE}
head(thedata2)
tddata2<-thedata2%>%
    gather(year, total, "1980":"2013")%>%
    filter(total != "..")

tddata2$total<- as.integer(tddata2$total)
    
tddata2<- tddata2%>%
    group_by(CntName,year)%>%
    summarise_each(funs(sum), total)%>%
    arrange(year)#%>%
    #filter(CntName %in% c("United States of America","Germany","France"))
#
#View(tddata2)
#tddata2 = NULL
```

####GGPlot
####1990 -1991 USSR Colapses, USSR Migration Explodes.
#### 2008 Black Market, Market Crashes, Germany migration Declines

```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=7}
require(plotly)
p <- ggplot(tddata2, aes(total/1000000, year, color = reorder(CntName, -total)))+
  geom_point(aes(size = total, frame = year, ids = CntName))+
  scale_fill_continuous()+
  labs(color = "Country Names", x = "Migration by Millions", y = "Years")

p <- ggplotly(p) %>%
    animation_opts(2000, easing = "elastic", redraw = TRUE)%>%
    add_annotations(
    yref="paper", 
    xref="paper", 
    y=1.1, 
    x=0, 
    text="Migration By Year/Country", 
    showarrow=F, 
    font=list(size=17)
  ) %>% 
  layout(title=FALSE)
p
# p = NULL

```


####Data2 posted by Ravi Itwaru (Airline Safety)
#### https://raw.githubusercontent.com/fivethirtyeight/data/master/airline-safety/airline-safety.csv

####Get data 3

```{r data3, message=FALSE, warning=FALSE}
theurl3<-"https://raw.githubusercontent.com/fivethirtyeight/data/master/airline-safety/airline-safety.csv"
thedata3 <- read.table(file = theurl3, header = TRUE, sep = ",") 
glimpse(thedata3)
```

####Tidy Data by Airline Incident Counts
Gather all incident totals and dipsplay GGPlot with Top 10 Airlines Incidents,then compare incidents to kilometers per week.

```{r message=FALSE, warning=FALSE}
head(thedata3)
tddata3<-thedata3%>%
    gather(incidents, count, 3:8)
    
tddata3$count<- as.integer(tddata3$count)

tddata3<-tddata3%>%
    filter(count > 0)%>%
    select(airline, count, avail_seat_km_per_week)%>%
    group_by(airline)%>%
    summarise_each(funs(sum), count, avail_seat_km_per_week)%>%
    arrange(desc(count))%>%
    top_n(10)
```

####GGPlot Top 10 Airport Incidents

```{r message=FALSE, warning=FALSE}
ggplot(data= tddata3) +
    geom_bar(mapping = aes(x = reorder(airline, -count), y = count, fill= airline), stat = "identity", position = "identity", show.legend = FALSE)+
theme(axis.text.x=element_text(angle=-90))

ggplot(data= tddata3) +
    geom_point(mapping = aes(x = avail_seat_km_per_week, y =count, color = airline, size = count))
 
#theme(axis.text.x = element_blank())
```
