---
title: "IS607 Project1"
author: "Anthony Pagan"
date: "September 23, 2018"
output:
  html_document:
    df_print: paged
---


####Import preformatted tmt2.csv file 


```{r}
library(stringr)
tst<-data.frame(read.csv(file = 'C:/temp/tmt2.csv', header = TRUE, sep= ",", stringsAsFactors = FALSE))
```

####Data preparation
Copy original dataframe to new list. Use gsub to partially format Ratings data. Remove front and end digits, spaces and miscellanous characters

```{r}
tst2<-tst

tst2$Rating<-str_trim(gsub(",>\\d*", "", gsub("^ \\d* ", "", gsub("\\d*/ R: ", "", tst$Rating))))
```

####Additional Data preparation
Copy dataframe to another dataframe. Clean up the remaing Rating unwanted strings that included characters, spaces and numbers. 

```{r}
tst3<-tst2
tst3$Rating<- str_trim(str_sub(tst2$Rating, start = 1 ,end =(str_locate(tst2$Rating, pattern = "(.?P. ?.)|( .)")[,1])))
```

####Clean and move Data
Insert data from tst2 to tst3 where tst3 ratings is NA

```{r}
for(i in 1:nrow(tst3))
{
    if (is.na(tst3$Rating[i]))
            {
               tst3$Rating[i] <-tst2$Rating[i]
            }
      
}
```

####Get Mean Data
Extracted numbers from each round by extracting digits only. Binded data to a dataframe and removed any data with NA. Took average of all rounds and added to a new column in tst2 dataframe.
```{r}
for(k in 1:nrow(tst3))
{
    k.w <-as.integer(str_extract(tst3[k,6:12], "\\d{1,2}"))
    df<- data.frame(x=numeric())
    for (m in k.w)
    {
      df<- rbind(df, data.frame(sum(as.integer(tst3[m,4], na.rm = TRUE))))
      sf<-sum(df, na.rm = TRUE)
      mn <- round(sf/na.omit(nrow(df)))
    }
    tst3$AvgPreChessRating[k]<-mn
}
```

####Update Columns
Updated column names with columns names listed in Project notes

```{r}
library(dplyr)

tst4 =rename(tst3, 
       "Player's Name" = "Player.Name", 
       "Player's State" = "State",
       "Total Number of Points" = "RoundTotal",
       "Player's Pre-Rating" = "Rating",
       "Average Pre Chess Ratings of Opponents" = "AvgPreChessRating")
```

####Display Data
Display a subset of data with only columns listed in Project notes.

```{r}
sdata<-subset(tst4, select = c(str_trim("Player's Name"), str_trim("Player's State"),str_trim("Total Number of Points"),
str_trim("Player's Pre-Rating"), "Average Pre Chess Ratings of Opponents"))
sdata
```

####Write data to CSV file

```{r}
write.csv(sdata, file="sdata.csv")
```
