---
title: "IS607 Assignement 5"
author: "Anthony Pagan"
date: "September 30, 2018"
output:
  html_document:
    df_print: paged
---

#
#
####Import CSV file file with stringsAsFactors set to false

```{r warning=FALSE}
td<-read.csv(file = 'tidydata.csv', stringsAsFactors = FALSE, sep = ",")
head(td)
```


####Setup to use Dplyr and Tidyr Libraries
```{r warning=FALSE}
##Load libraries
library(dplyr)
library(tidyr)
```


#### Remove any rows with NA
```{r warning=FALSE}
#Omit NA rows
td1 <-na.omit(td)
td1
```

####Update table with any missing Airline names
```{r warning=FALSE}
td1$X<- c("ALASKA", "ALASKA", "AM WEST", "AM WEST")
```

####Gather City Times into 1 column
```{r warning=FALSE}
td2<-td1%>%
    gather(city, value, 3:7)
td2
```

#####Spread the X.1 column to 2 columns for Delayed and OnTime data
```{r warning=FALSE}
td3<-spread(td2,X.1, value)
td3
```

####Rename columns with descriptive names
```{r warning=FALSE}
td3<-rename(td3, "airline"="X",
            "on-time" = "on time")
td3
```

####Analysis: Average delay by airline shows AM WEST had the highest average delay
```{r warning=FALSE}
avgdlbyal<-td3 %>% group_by(airline)%>% summarise(delayedmean = mean(delayed))%>%arrange(airline, delayedmean)
avgdlbyal
```

####Analysis: Average delay by city shows Phoenix and Seattle had the highest average delays
```{r warning=FALSE}
avgdlbyct<-td3 %>% group_by(airline,city)%>% summarise(delayedmean = mean(delayed))%>%arrange(airline, delayedmean)
avgdlbyct
```

####Graph confirms 2 Cities with highest rate is Seatel and Pheonix and AM West inclued the city wiht  highestaverage delays
```{r warning=FALSE}
library(ggplot2)
ggplot(data= td3, mapping = aes(x = airline, y=delayed)) +
    geom_point(mapping = aes(size = delayed,  color = city))
```

