---
title: "Interactive Plots"
author: "Anthony Pagan"
date: "October 28, 2018"
output: 
  slidy_presentation: 
    keep_md: yes
    smaller: yes
    widescreen: yes
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

## Introduction - Buiding Interactive Plots in R
    - Using Plotly
    - Plotly Examples
    - Using Shiny App
    - Shiny Examples
    - Shiny with Plotly Example
    - Recap
    - Questions

## Plotly Overview

- What is Plotly

    - An open source javascript library
    - Allows you to create web based interactive Graphs
    - Intial goal was to suport 3D and Mesh chart that types ggplot2 did not support. 
    -plotly uses the htmlwidget framework, which allows plots to work seamlessly and consistently in various contexts  
    - Works with the most basic charts to advance statistical or scientific charts
    - Can be use with local browser or published to a ploty web service

## Shiny Overview
    
- What is Shiny

    - Also interactive same as Plotly
    - Shiny can be used as app and can be servered directly from the web
    - Ploty can be in conjuction with Shiny app.
    
    

## Using Plotly

- URL

    - Documentation:  https://plot.ly/r/
    - Online Book: https://plotly-book.cpsievert.me/index.html
    - Cheatsheet: https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf?_ga=2.137841968.128055842.1542046850-981781115.1542046850

- Library

    - install.packages("plotly") or
    
    - install the latest development version (on Github) via devtools:devtools::install_github("ropensci/plotly")
    
    - 2 ways to initiate:
            - plot_ly() function for datasets
            - ggplotly() function totransforms a ggplot object into a plotly                  - ggplotly has a dependency on ggplot2 while plot_ly does not.

## Basic Example 1

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(plotly)
library(ggplot2)
plot_ly(z = ~volcano)

```


## Basic Example 2

```{r echo=TRUE, message=FALSE, warning=FALSE}

#options(browser = 'FALSE')
t<-plot_ly(midwest, x = ~percollege, color = ~state, 
type = "box",width = 800, height = 300)%>%
layout(margin = list(l = 100))
t
rangeslider(t)
```

## Using Plotly with GGPlotly: GGPLOT  

```{r echo= TRUE, message=FALSE, warning=FALSE}

p <- ggplot(txhousing, aes(date, median,
    width = 800, height = 300)) +
    geom_line(aes(group=city), alpha = 0.2)
p

```


## Using Plotly with GGPlotly

```{r echo= TRUE, message=FALSE, warning=FALSE}

#options(browser = 'FALSE')
subplot(
    p, ggplotly(p, tooltip = "city",
                width = 800, height = 300),
    ggplot(txhousing, aes(date, median)) + geom_bin2d(),
    ggplot(txhousing, aes(date, median)) + geom_hex(),
    nrows =2, shareX= TRUE, shareY = TRUE,
    titleY = FALSE, titleX = FALSE
    )

```
   

## Statistical Example


```{r echo= TRUE, message=FALSE, warning=FALSE}
library(plotly)
library(GGally)
pm<- ggpairs(iris)
ggplotly(pm,width = 800, height = 300)

```

## Plotly online example

    - Setup Account in https://plot.ly
    - Get username and secret api key from https://plot.ly/settings/api#/
    - Use Sys.setenv to setup username and passowrd keys
    - Api_create to upload file.

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
Sys.setenv("plotly_username"="apag101")
Sys.setenv("plotly_api_key"="OyQslu5MKVQcEjx7aVJV")
api_create(t, filename = c("test"))

```

## Using Plotly Web Figures

    - Use api_dowload(id, username) to download plot from site
    - Example of dowloading from my account and another public account.
    

```{r echo=TRUE, message=FALSE, warning=FALSE}
#Download from current account
api_download_plot("8","apag101")

#Download from https://plot.ly/~cpsievert/559/_1-vs-2/#/
api_download_plot("559", "cpsievert")

```

## Using Shiny
    
    - URL
    
        - Documentation:  https://www.rstudio.com/products/shiny/
        - Online Book and Videoes: http://shiny.rstudio.com/tutorial/
        - Cheatsheet: https://www.rstudio.com/resources/cheatsheets/#shiny

- Library

    - install.packages("shiny") 
    - install.packages('rsconnect')

## App R Shiny Example

    - Uses app.R file to run
    - app.R has three components:
        - UI is theuser interface object
        - server function has to instructions to run build the app
        - shinyApp function takes the UI and Server object and creates shiny app

    
```{r message=FALSE, warning=FALSE, include=FALSE}
library(shiny)
# Define UI for application that draws a histogram
ui <- fluidPage(
   # Application title
   titlePanel("Old Faithful Geyser Data"),
   # Sidebar with a slider input for number of bins 
   sidebarLayout(
      sidebarPanel(
         sliderInput("bins",
                     "Number of bins:",
                     min = 1,
                     max = 50,
                     value = 30)
      ),
      
      # Show a plot of the generated distribution
      mainPanel(
         plotOutput("distPlot")
      )
   )
)

# Define server logic required to draw a histogram
server <- function(input, output) {
   
   output$distPlot <- renderPlot({
      # generate bins based on input$bins from ui.R
      x    <- faithful[, 2] 
      bins <- seq(min(x), max(x), length.out = input$bins + 1)
      
      # draw the histogram with the specified number of bins
      hist(x, breaks = bins, col = 'darkgray', border = 'white')
   })
}

# Run the application 
shinyApp(ui = ui, server = server)
```

## Shiny Online Example

    - Setup a shiny account https://www.shinyapps.io
    - Get username and user key from https://www.shinyapps.io/admin/#/dashboard
    - install rconnect library(rsconnect)
    - Deploy web app: rsconnect::deployApp('path/to/your/app')

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#install.packages('rsconnect')

rsconnect::setAccountInfo(name='apag101', token='98AFD680E6AB821CF1F430EF2A6DABFC', secret='rgMjLtYC8JcXgTFRb5TWOSzKlLXk4W4Nkp4fr1eU')

library(rsconnect)
rsconnect::deployApp('C:/Users/apagan/Documents/test2')

```

## Shiny App With Plotly 

    - Render Ploty plot in shiny app

    
```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

ui2 <- fluidPage(
  plotlyOutput("plot"),
  verbatimTextOutput("event")
)

server2 <- function(input, output) {

  # renderPlotly() also understands ggplot2 objects!
  output$plot <- renderPlotly({
    plot_ly(mtcars, x = ~mpg, y = ~wt)
  })

  output$event <- renderPrint({
    d <- event_data("plotly_hover")
    if (is.null(d)) "Hover on a point!" else d
  })
}

shinyApp(ui2, server2)

```

## Recap




## QUESTIONS?